Initial Analysis

### 2019-08-21 Initial_Analysis.Rmd
Updated analuysis and did some results exploration (de-bugging). Updated function to output all models and non transboundary species

# Set paths depending on machine Beas or Hall1000
if(Sys.info()[7] == "jepa88"){
  Data_Path <- "Z:/JULIANO_NEYMAR/FishForVisa/"
  Path <- paste(Data_Path,"Results/",sep = "")
  }
if(Sys.info()[7] == "carmelia"){
  Data_Path <- "~/GitHub/FishForVisa/Temporal_Data/"
  Path <- paste(Data_Path,sep = "")
}else{
  Data_Path <- "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/"
  Path <- paste(Data_Path,"Results/",sep = "")
}

@@

@@

#### For testing ###
# Spp <- 600004
# Model <- "All"
# Neighbors <- Neighbor_List
# Coord = CoorG
# __________________________________

@@

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbors$INDEX,# Filter data only located within EEZs
           Model != "SAU_C") %>% # Only using observational and modelled data
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = mean(Value,na.rm=T)
              ) %>%
    filter(Model_Index > 0) %>%
    #____________ ESTIMATING FUNDAMENTAL NICHE (TRESHOLD 2)_________ #
    left_join(SppDist,
              by = c("TaxonKey","INDEX")) %>% 
    filter(Model == "SAU_C", # Only keeping cells where SAU catch exists
           Value > 0) %>% 
    mutate(Model_Index = Model_Index*100) %>% 
    select(-Model, -Value, -Latitude,-Longitude) %>% 
    left_join(EEZ_CellID,
              by = "INDEX") %>% 
    left_join(EEZIDs_List,
              by = "EEZID")

  # MODEL INDEX (TRESHOLDS 1 & 2) #
  Model_Index_D <- Trans_Spp %>% 
    group_by(Name,
             TaxonKey,
             Model_Index
    ) %>% 
    summarise(n_cells_spp = n()) %>% 
    select(-n_cells_spp)

@@

  # Step 2. Determines the amount of grids present in each country
  Spp_Grid <- Trans_Spp %>% 
    group_by(TaxonKey,Model_Index,EEZID) %>% 
    summarise(n_spp_eez = length(unique(INDEX))) %>% 
    left_join(Neighbors_EEZ,
              by = "EEZID") %>% 
    filter(Country_Territory %in% Trans_Spp$Name, #Filter out unwanted neighbors (those who don't have grids within but get included because they are neighbors)
           Neighbor_Territory %in% Trans_Spp$Name)

@@

  # Split dataframes to merge latter
  Country_Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(Model_Index,TaxonKey,Name=Country_Territory,n_spp_eez)

  Neighbor_Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(Model_Index,TaxonKey,n_spp_eez,Name=Neighbor_Territory, Country_Territory)

  # Merge dataframes to get totals per neighbords
  Area_Index_D <- full_join(Country_Territory_T,
                           Neighbor_Territory_T, by = c("Model_Index","Name","TaxonKey")) %>% 
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per neighbors
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(Neighbor_Territory = Name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbor = n_spp_eez.y) %>% 
    mutate(Area_Index = n_spp_Country/Spp_Total) %>%  #Estimates the proportion of grids per country
    select(1:3,6,8)

  #____________ FINAL INDEX TABLE_________ #
  Indexes <- Model_Index_D %>% 
  rename(Country_Territory = Name) %>% 
  full_join(Area_Index_D,
            by = c("TaxonKey","Country_Territory","Model_Index")
  ) %>% 
    filter(!is.na(Area_Index),
           !is.na(Model_Index)
           )%>%  # Remove cases where the country does not pass the area_index and viceversa
    select(TaxonKey,Country_Territory,Neighbor_Territory,everything())

@@

@@

# For Function
Exploited_Species_List <- Exploited_Species$TaxonKey

@@

@@

# Details

## First world Anchovies

@@

head(Results_Data)
unique(Results_Data$Model_Index)
length(unique(Results_Data$TaxonKey)) #672 spp

Clean_Results <- Results_Data

### Problem solving

@@

Neighbord_Territory_T <- Results_Data %>% 
  group_by(eez_name=Neighbor_Territory) %>% 
  summarise() %>%
  left_join(sau_eez_fes,
            by = "eez_name") %>% 
  select(Neighbor_Territory= eez_name,
         NT_fishing_entity=fishing_entity
         )

Country_Territory_T <- Results_Data %>% 
  group_by(eez_name=Country_Territory) %>% 
  summarise() %>%
  left_join(sau_eez_fes,
            by = "eez_name")  %>% 
  select(Country_Territory= eez_name,
         CT_fishing_entity=fishing_entity)

Clean_Results <- Results_Data %>%   
  left_join(Neighbord_Territory_T,
              by ="Neighbor_Territory") %>% 
  left_join(Country_Territory_T,
              by ="Country_Territory") %>% 
  mutate(
    Same_Nation = ifelse(CT_fishing_entity==NT_fishing_entity,"Yes","No")
  ) %>% 
  filter(Same_Nation == "No")

TenTwenyfive <- Clean_Results %>% 
  filter(Model_Index == 100,
         Area_Index >= 0.25,
         Area_Index <= 0.75
         ) %>% 
  group_by(Name=Country_Territory,Model_Index) %>% 
  summarise(
    n_trans = length(unique(TaxonKey)),
    n_countries = length(unique(Neighbor_Territory)),
    Trans_Rate = n_trans/n_countries
    ) %>% 
  filter(Model_Index == 100)
  # head(Results_Data)

# The path
path_world <- paste("~/Documents/UBC/Oceans_Project/Distribution/Data/Spatial/World_EEZ_v9_20161021",sep="")
# #The File
fnam_world <- "eez_boundaries.shp"
# #Load it!
# 
World_EEZs <- st_read(dsn = path_world,
                      layer =file_path_sans_ext(fnam_world))
# 
head(World_EEZs)

# head(World_EEZs_simple)

pal <- wes_palette("Zissou1", 100, type = "continuous")

Plot <- ggplot(World_EEZs_simple) +
  geom_sf(aes(fill = n_trans)) +
  scale_fill_gradientn(colours = pal) +
  ggtheme_map()

>>>>

# Set paths depending on machine Beast (jepa88), "carmelia" or Hall1000
if(Sys.info()[7] == "jepa88"){
  Data_Path <- "Z:/JULIANO_NEYMAR/FishForVisa/"
  Path <- paste(Data_Path,"Results/All_Models/",sep = "")
  }
if(Sys.info()[7] == "carmelia"){
  Data_Path <- "~/GitHub/FishForVisa/Temporal_Data/"
  Path <- paste(Data_Path,sep = "")
}else{
  Data_Path <- "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/"
  Path <- paste(Data_Path,"Results/",sep = "")
}

@@

** NOTE THAT THE FUNCTION IS SETTED TO OUTPUT RESULTS FOR LL MODELS **

@@

#### For testing 
# Spp <- 600059
# Model <- "All"
# Neighbors <- Neighbor_List
# Coord = CoorG
# __________________________________

@@

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbors$INDEX,# Filter data only located within EEZs
           Model != "SAU_C") %>% # Only using observational and modelled data
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  
    filter(Value > 0) %>%  # Remove latter, just keeps places with 1
    # Commented out to produce a map for each dataset
    # group_by(TaxonKey,
    #          INDEX
    # ) %>%
    # summarise(Model_Index = mean(Value,na.rm=T)
    #           ) %>%
    # filter(Model_Index > 0) %>%
    #____________ ESTIMATING FUNDAMENTAL NICHE (TRESHOLD 2)_________ #
    # left_join(SppDist,
    #           by = c("TaxonKey","INDEX")) %>% 
    # filter(Model == "SAU_C", # Only keeping cells where SAU catch exists
    #        Value > 0) %>%
    # mutate(Model_Index = Model_Index*100) %>% 
    # select(-Model, -Value, -Latitude,-Longitude) %>% 
    left_join(EEZ_CellID,
              by = "INDEX") %>% 
    left_join(EEZIDs_List,
              by = "EEZID")

  # MODEL INDEX (TRESHOLDS 1 & 2) #
  Model_Index_D <- Trans_Spp %>% 
    group_by(Name,
             TaxonKey,
             # Model_Index, # Un-comment after producing models x datasets
             Model # Remove after producing models x datasets
    ) %>% 
    summarise(n_cells_spp = n()) %>% 
    select(-n_cells_spp)

@@

  # Step 2. Determines the amount of grids present in each country
  Spp_Grid <- Trans_Spp %>% 
    group_by(TaxonKey,
             # Model_Index, # Un-comment after producing models x datasets
             Model, # Remove after producing models x datasets
             EEZID) %>% 
    summarise(n_spp_eez = length(unique(INDEX))) %>% 
    left_join(Neighbors_EEZ,
              by = "EEZID") %>% 
    filter(Country_Territory %in% Trans_Spp$Name, #Filter out unwanted neighbors (those who don't have grids within but get included because they are neighbors)
           Neighbor_Territory %in% Trans_Spp$Name)

@@

  # Split dataframes to merge latter
  Country_Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(
      # Model_Index, # Un-comment after producing models x datasets
      Model, # Remove after producing models x datasets
      TaxonKey,Name=Country_Territory,n_spp_eez
    )

  Neighbor_Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(
      # Model_Index,
      Model, # Remove after producing models x datasets
      TaxonKey,n_spp_eez,Name=Neighbor_Territory, Country_Territory
    )

  # Merge dataframes to get totals per neighbords
  Area_Index_D <- full_join(Country_Territory_T,
                           Neighbor_Territory_T, by = c(#"Model_Index",
                             "Name","TaxonKey","Model")) %>% # Remove Model abd undomment Model_Index after producing Models x datasets
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per neighbors
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(Neighbor_Territory = Name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbor = n_spp_eez.y) %>% 
    mutate(Area_Index = n_spp_Country/Spp_Total) %>%  #Estimates the proportion of grids per country
    # select(1:3,6,8) # Commented out for all models analysis
    select(1:4,6,8)

  #____________ FINAL INDEX TABLE_________ #
  Indexes <- Model_Index_D %>% 
  rename(Country_Territory = Name) %>% 
  full_join(Area_Index_D,
            by = c("TaxonKey","Country_Territory",
                   #"Model_Index"
                   "Model") # Remove Model abd undomment Model_Index after producing Models x datasets
  ) %>% 
    filter(!is.na(Area_Index)#,
           # !is.na(Model_Index) # Uncomment after producing Models x datasets
           )%>%  # Remove cases where the country does not pass the area_index and viceversa
    select(Model,TaxonKey,Country_Territory,Neighbor_Territory,everything())

@@

  if(nrow(Indexes) == 0 & Save == "Y"){

# Reads dataset of Non transboundary species
    Non_Transboundary_Data <- fread(paste(Results_Path,"Non_Transboundary_Data.csv",sep=""))

    # Get current Spp that is not transboundary 
    Current_Spp <- tibble(Non_trans = Spp)

    # Combine new spp with others
    New_Non <- Non_Transboundary_Data %>% 
      bind_rows(Current_Spp)

    # Set a name for file and saves data
    New_Non_Name <- paste(Results_Path,"Non_Transboundary_Data.csv",sep="")
    write.csv(New_Non, New_Non_Name, row.names = F)

@@

# For Function
Exploited_Species_List <- Exploited_Species$TaxonKey[1:3]

@@

# Creates dummy dataset for Non transboundary species
write.csv(tibble(Non_trans = "NA"), paste(Results_Path,"Non_Transboundary_Data.csv",sep=""), row.names = F)

@@

# Data needed to remove duplicates
sau_eez_fes <- read.csv(paste(Data_Path,"Spatial_Data/sau_eez_fes.csv",sep=""))

Clean_sau_eez_fes <- sau_eez_fes %>% 
  gather("Type","FishingRegion",2,4) %>% 
  left_join(sau_eez_fes,
            by = ("fishing_entity_id")
            ) %>% 
  group_by(FishingRegion,fishing_entity) %>% 
  summarise(n())

# head(Results_Data)
# unique(Results_Data$Model_Index)
length(unique(Results_Data$TaxonKey)) #672 spp

# Clean_Results <- Results_Data

# The path
path_world <- paste(Data_Path,"Spatial_Data/SAU_Shapefile/",sep="")
# #The File
fnam_world <- "SAUEEZ_July2015.shp"
# #Load it!
# 
World_EEZs <- st_read(dsn = path_world,
                      layer =file_path_sans_ext(fnam_world))
# 
# head(World_EEZs)

Clean_Results <- Indexes %>%   
  rename(FishingRegion=Country_Territory) %>% 
  left_join(Clean_sau_eez_fes,
            by = "FishingRegion") %>% 
  select(Model,
         TaxonKey,
         Country_Territory= FishingRegion,
         FishingRegion=Neighbor_Territory,
         Area_Index,
         CT_fishing_entity= fishing_entity) %>% 
  left_join(Clean_sau_eez_fes,
            by = "FishingRegion") %>% 
  select(#1,2, 
         Neighbor_Territory = FishingRegion,
         #Model_Index:CT_fishing_entity,
         NT_fishing_entity= fishing_entity,
         everything()
  ) %>% 
  mutate(
    CT_fishing_entity = ifelse(is.na(CT_fishing_entity),paste(Country_Territory),paste(CT_fishing_entity)),
    NT_fishing_entity = ifelse(is.na(NT_fishing_entity),paste(Neighbor_Territory),paste(NT_fishing_entity)),
    Same_Nation = ifelse(CT_fishing_entity==NT_fishing_entity,"Yes","No")
  )

# Manually deal with NAs... Ughhhhh
Results_NA <- Clean_Results %>% 
  filter(is.na(Same_Nation)) %>% 
  group_by(Country_Territory,Neighbor_Territory) %>% 
  summarise(n())

TenTwenyfive <- Clean_Results %>% 
  filter(#Model_Index == 100,
         Area_Index >= 0.25,
         Area_Index <= 0.75,
         Same_Nation == "No"
         ) %>% 
  group_by(Name=Country_Territory,Model #,Model_Index
           ) %>% 
  summarise(
    n_trans = length(unique(TaxonKey)),
    n_countries = length(unique(Neighbor_Territory)),
    Trans_Rate = n_trans/n_countries
    ) #%>% 
  # filter(Model_Index == 100)
  # head(Results_Data)

# head(World_EEZs_simple)
# World_EEZs_simple %>% 
#   filter(Name == "Chile")

pal <- wes_palette("Zissou1", 1000, type = "continuous")

Plot <- ggplot(World_EEZs_simple) +
  geom_sf(aes(fill = n_trans)) +
  scale_fill_gradientn(colours = pal,
                       na.value = "white") +
  ggtheme_map() +
  facet_wrap(~Model)

# Problem solving

### First world Anchovies

@@

@@

#### Missing stocks

There are places like Peru-Chile or Alaska-Canada that should have trans spp... In Alaska the issue is that Alaska doesn't actually exist as a separate shapefile...

```{r Missing_stocks, eval =F}

Exploited_Species <- fread(paste(Data_Path,"Distribution_Data/exploited_species_list.csv",sep=""))

Clean_sau_eez_fes <- sau_eez_fes %>% 
  gather("Type","FishingRegion",2,4) %>% 
  left_join(sau_eez_fes,
            by = ("fishing_entity_id")
            ) %>% 
  group_by(FishingRegion,fishing_entity) %>% 
  summarise(n())

# The Alaska paradox
# It's a problem of the NAs, fixed.

USA_West <- Results_Data %>% 
  filter(Country_Territory == "USA (West Coast)") %>% 
  left_join(Exploited_Species,
            by = "TaxonKey")

# Northern sea Russia
Russia_map <- World_EEZs %>% 
  filter(stringr::str_detect(Name,"Russia")) %>% 
  st_simplify(preserveTopology = TRUE, dTolerance = 100)

head(Russia_map)

ggplot(Russia_map) +
  geom_sf(aes(fill = Name)) 
# Shapefiles not matching
# Russia (Laptev to Chukchi Sea)
# Russia (Kara Sea)

Russian_Regions <- c("Russia (Laptev to Chukchi Sea)","Russia (Kara Sea)")

Russia_Results <- Results_Data %>% 
  filter(stringr::str_detect(Country_Territory,"Russia"))

unique(Russia_Results$Country_Territory)
# Missing Laptev to Chukchi sea

### What about Vicky's data
#SAU / Vicky's data
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))

Russia_EEZlist <- EEZIDs_List %>% 
  filter(stringr::str_detect(Name,"Russia"))

unique(Russia_EEZlist$Name)

### So maybe there are no neighbors?
Neighbor_List <- fread(paste(Data_Path,"Spatial_Data/Neighbor_List.csv", sep =""))

Russia_Neighbor <- Neighbor_List %>% 
  filter(stringr::str_detect(Name,"Russia"))

unique(Russia_Neighbor$Country_Territory)

# OK So Kara Sea has no borders outside Russia, hence white. Now Laptev does have borders but No species... which is weird... Paific Halibut?

# I manually ran the function to see what's going on

# ggplot() +
#   geom_sf(data = Russia_map, aes(colour = Name)) +
#   geom_tile(data = subset(SppDist, Value !=0),
#             aes(
#     x = Longitude,
#     y = Latitude,
#     fill = log10(Value)
#   ) 
#   ) +
#   facet_wrap(~Model)

# The main issue is that there is no catch data there...
